---
title: "lab_3_part_1_hgw"
author: "Hannah Garcia-Wickstrum"
date: "1/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(GGally)
library(jtools)
```

```{r}
penguins %>% 
  select(species, bill_length_mm:sex) %>% 
  ggpairs(aes(color=species))
```

create a subset with only Adelie and Chinstrap penguins
```{r}
adelie_chinstrap <- penguins %>% 
  filter(species %in% c("Adelie", "Chinstrap")) %>% 
  mutate(species = fct_drop(species)) # dropped gentoo as one of the species
```

Let's just do a data viz of those two species:
```{r}
ggplot(data=adelie_chinstrap, aes(x=body_mass_g, y = flipper_length_mm)) +
  geom_point(aes(color=sex)) +
  facet_wrap(~species)
```

Binary logistic regression 

- body mass
- flipper length
- penguin sex 

```{r}
adelie_chinstrap_blr <- glm(species ~ body_mass_g + flipper_length_mm + sex,
                            data = adelie_chinstrap,
                            family = "binomial")
```

Take a look at the outcomes:
```{r}
adelie_chinstrap_blr

summary(adelie_chinstrap_blr)
# adelie is the first species listed and so it is the reference species

# Flipper length = 1.778e-01 (the odds of being a chinstrap penguin is increasing) as flipper length increases, the odds of that species being a chinstrap is increasing
```

```{r}
ggplot(data = adelie_chinstrap, aes(x = species, y = flipper_length_mm)) +
  geom_jitter(aes(color=sex))

# maybe chinstraps tend to be a bit larger than adelie penguins. As flipper length gets longer, then it's more likely that you are a chinstrap penguin. That positive coefficient from above does align with what we are seeing here. 
```

What are the actual probabilities of being Chinstrap for each of the existing observations in our adelie_chinstrap data frame?

```{r}
blr_fitted <- adelie_chinstrap_blr %>% 
  broom::augment(type.predict = "response")
```

Plotting fitted probabilities:
```{r}
ggplot(data = blr_fitted, aes(x=flipper_length_mm, y = .fitted)) +
  geom_point(aes(color=sex, pch = species)) +
  geom_smooth(aes(color=sex), se=FALSE) +
  labs(x = "Flipper length (mm)",
       y = "Probability of outcome Chinstrap")
```

Plot the model based on different variables:
```{r}
effect_plot(adelie_chinstrap_blr,
            pred = flipper_length_mm,
            y.label = "Probability of Chinstrap",
            interval = TRUE)

effect_plot(adelie_chinstrap_blr,
            pred = body_mass_g,
            y.label = "Probability of Chinstrap",
            interval = TRUE)
```


What's the probability that a female penguin weighing 3410 g with a flipper length 192mm is a Chinstrap?

```{r}
ex_1 <- predict(adelie_chinstrap_blr,
                data.frame(sex = "female",
                           body_mass_g = 3410,
                           flipper_length_mm = 192), # just doing this will provide log odds
                type = "response") # adding this will provide probability
ex_1
```




